---
title: "corpusr"
author: "Charlie Gallagher"
date: "11/24/2020"
output: 
  rmarkdown::github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This package, called `corpusr`, is a simple, very high-level package for querying books from `gutenbergr` for sentences containing certain words or regular expressions. 

# Quickstart
 Before starting, you should have both `corpusr` and `gutenbergr` installed. When `corpusr` is loaded, so is `gutenbergr`.

```{r}
library(tidyverse)
library(corpusr)
```

### Querying a Work
The simplest search you can do is a single work. For this, use `corpusr::guten_search()`, which takes the arguments "code" and "word". The code is a number that comes from `gutenbergr::gutenberg_works()`, which is a dataset containing an observation for every work in Project Gutenberg. Search this for the work you want to analyze and get its `gutenberg_id`. For example, here is how to search the US Constitution for the word "Enumeration."

```{r}
gutenbergr::gutenberg_works(title == "The United States Constitution") %>% 
  pull(gutenberg_id) %>% 
  guten_search("Enumeration")
```

The search tools are written with regular expressions built in, so you can search for several word forms at once, sentences, or common collocations. The regular expression is automatically book-ended by word separators, so 'the' will never match 'there', for example. 

```{r}
gutenbergr::gutenberg_works(title == "The United States Constitution") %>% 
  pull(gutenberg_id) %>% 
  guten_search("House of Representatives")
```


When more than 10 sentences are returned for a single work, the output is truncated. 

```{r}
gutenbergr::gutenberg_works(title == "The United States Constitution") %>% 
  pull(gutenberg_id) %>% 
  guten_search("the")
```

### Querying Multiple Works
You can search multiple works if you provide a vector of `gutenberg_id`s with `corpusr::guten_search_works()`. Here, I search for any books with "Puget Sound" in their title for the words "serene" or "serenely". 

```{r}
gutenbergr::gutenberg_works(str_detect(title, "Puget Sound")) %>% 
  pull(gutenberg_id) %>% 
  guten_search_works("serene(ly)?")
```

In fact, five books were returned, but only those containing 'serene' or 'serenely' are included in the final `cps_corpus` object. (A `cps_corpus` is a list of `cps_bk` objects. More on the `cps` classes later, though.)

### Querying an Author's Works
Commonly, one wants to search all an author's works for a certain word or phrase. For this, you can query the `gutenbergr` database as above or use `corpusr::guten_search_author()`. 

```{r}
corpusr::guten_search_author("Lincoln, Abraham", "(S|s)cores?")
```

Twelve works by Abraham Lincoln are returned (with some odd entries owing to how the sentence parser works with tables of contents and the like), with all the sentences that contain "score" or "scores". 

When you enter the name of an author, it must be the exact form that it appears in in the `gutenbergr` dataset. Thus, you might still prefer to use `stringr::str_detect` and `gutenbergr::gutenberg_works` when you don't know that exact form. 


### Building a Corpus
Functions dealing with corpora have the prefix `corpus_*`. To build a corpus from the `gutenbergr` database, you can use the `corpusr` functions `corpus_build_works` or `corpus_build_author`. These return a list of complete `cps_bk` objects, each of which contains the full texts with sentence-level observations, as opposed to the _filtered_ `cps_bk` objects returned by `guten_search`, for example. Depending on the function you use, you provide either a vector of work IDs (`corpus_build_works`) or an author name. Currently, you cannot pass a vector of author names, although this functionality will be added in future versions. 

```{r}
corpus_build_works(4) %>% print()
```


This corpus is small, but good for illustrating the idea. Once built, you can search this corpus using `corpusr::corpus_search()`:

```{r}
corpus_build_works(4) %>% 
  corpus_search("battlefield")
```

In general, it's best work from a corpus. A corpus has several advantages over the interactive `guten_*` functions:

1. A corpus built with `corpus_build_*` can be saved, allowing you to 
2. You are guaranteed to always be working with the same data, regardless of whether Project Gutenberg changes the files.
3. It's much faster to use `corpus_search` than it is to use `guten_search_works`. 


# The `corpusr` Package
## Motivation
The `gutenbergr` package allows one to query the Project Gutenberg book database for searchable full-texts, which you can then analyze in R. What I wanted to do was search for illustrative examples of sentences in these works. For this purpose, there are several weaknesses in the `gutenbergr` package that I wanted to improve on:

1. **Lines, not sentences** The responses are organized only into data frames with a 'full-text' column, in which the unit of observation is the line. 
2. **Lack of grouping structures** Querying multiple books can be difficult because there is no structure that allows you to easily distinguish books and authors from each other. 
3. **No print methods** Without a convenient print method, printing many sentences from many books becomes an exercise in data manipulation. 

In `corpusr`, the unit of observation is the sentence, and books (`cps_bk` objects) can be grouped together into a `cps_corpus` object, which provides very clear grouping and printing. 

## Implementation
There are two objects, `cps_bk` and `cps_corpus`. A `cps_bk` stores a list containing the full text (as sentences, not lines), the author, and the title of the work. This way, unnecessary metadata is kept to a minimum. The print method for a `cps_bk` is designed to make the author and title clear. For example:

```
Author:  Lincoln, Abraham 
Title:  Noted Speeches of Abraham Lincoln
Including the Lincoln-Douglas Debate 

1. Occasional poisonings from the kitchen and open or stealthy assassinations in the field, and local revolts extending to a score or so, will continue to occur as the natural results of slavery; but no general insurrections of slaves, as I think, can happen in this country for a long time. 
```

A list of `cps_bk` objects can be made into a `cps_corpus`, which provides a further print method that makes several things clear:

1. The author, title, and number of returned sentences, via the `cps_bk` print method.
2. The number of books containing the searched-for word or phrase. 

For example: 

```
[[1]]

Author:  Lincoln, Abraham 
Title:  Lincoln's Gettysburg Address
Given November 19, 1863 on the battlefield near Gettysburg, Pennsylvania, USA 

1. Lincoln's Gettysburg Address, given November 19, 1863 on the battlefield near Gettysburg, Pennsylvania, USA   Four score and seven years ago, our fathers brought forth upon this continent a new nation:  conceived in liberty, and dedicated to the proposition that all men are created equal.  


-----------------------------------------------------

< ... TRUNCATED (not part of print method) ... >

[[11]]

Author:  Lincoln, Abraham 
Title:  Speeches & Letters of Abraham Lincoln, 1832-1865 

1. He says that, upon the score of equality, slaves should be allowed to go into a new Territory like other property. 

2. There are scores of them--good men in their character for intelligence, for talent and integrity. 

3. Occasional poisonings from the kitchen, and open or stealthy assassinations in the field, and local revolts extending to a score or so, will continue to occur as the natural results of slavery; but no general insurrection of slaves, as I think, can happen in this country for a long time. 


-----------------------------------------------------

[[12]]

Author:  Lincoln, Abraham 
Title:  Noted Speeches of Abraham Lincoln
Including the Lincoln-Douglas Debate 

1. Occasional poisonings from the kitchen and open or stealthy assassinations in the field, and local revolts extending to a score or so, will continue to occur as the natural results of slavery; but no general insurrections of slaves, as I think, can happen in this country for a long time. 


-----------------------------------------------------
```

At the moment, the regular expression is not saved or printed, but this is on the to-do list. 



---

This is one project I would like to convert to a Python project using the [Natural Language Toolkit](https://www.nltk.org/). Owing to the large size of the corpora that you can generate with `corpusr`, both packages would have a place in my language processing workflow. But this is all just personal notetaking.

---

# History and Purpose
I read often, and I write often. When I find a word I like, say 'jeremiad', I'm interested in how the word is used. A 'jeremiad' is a lamentation or complaint, and it refers specifically to the Jeremiah in the Bible: does that mean it's always used with a religious connotation? Does it have a typical collocation?

Illustrative quotations are central to modern lexicography. The first edition of the OED contained 252,200 entries and 1,861,200 illustrative quotations from a corpus of approximately 5 million slips, all hand-written and contributed by everyone lexicographical authorities to interested readers. Today, the OED contains about 290,500 main entries with 2,412,400 illustrative quotations. (For an interesting look at the statistics of the OED, take a look at the [OED Editions](https://public.oed.com/history/oed-editions/) website.) There was, is, and always will be, a large space for quotations in the OED, because lexicographers are as interested in word usage as in word definitions (if it even makes sense to talk about them as separate entities). Today, the OED is powered by the Oxford English Corpus. It contains nearly 2.1 billion words, all searchable in context, from all imaginable sources from the English-speaking world.

Important developments often come from extremes. The OED citations were often from works of literature and other carefully edited works. In 1961, Merriam-Webster published _Webster's Third New International Dictionary_, and the editors (led by Philip Gove) included all manner of quotations from well-respected literary men to the Saturday Evening Post and Reader's Digest. Reviewers were outraged. They felt that proper English was being degraded, destroyed. But to many lexicographers it made sense. It certainly made sense to the editors of the OED. The Oxford English Corpus contains a much wider array of citations from outside literature. 

The purpose of my package is small, but it derives mostly from one thing: it's hard to find good illustrative quotations. Most corpora are not publicly available, and if they are they are expensive, difficult to use, or small. I wanted a search engine in which you could search for a word or phrase and be given example sentences from across literature. 

A side note: This package has a focus on literature, and literature does not represent the language as it is spoken; neither, then, does this package. The `corpusr` package is useful for studying the standard written English contained in books whose copyrights have expired and left them in the public domain. 

Why is it necessary to look at example sentences? Why not use a good dictionary with usage labels like _rare_, _chiefly N. American_, and so on? There is no reason not to use a dictionary, but it's analogous to reading a description of a photograph instead of looking at the photograph itself. By reading a description of a photo, you understand the content, the context, and you might get more information in a description than you could possibly by looking at the photograph (its history, inspiration, etc.). But you don't understand the photograph because you haven't seen it.

Today, there are no dictionaries of illustrative quotations. Even the OED gives mainly historical quotations, not all of which are useful to the modern English speaker/writer. Thus, there is a need to have a simple-to-use, searchable, customizable corpus of sentences.
