---
title: "Corpusr: Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Corpusr: Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(corpusr)
```

## Introduction
The `corpusr` package shows you words in context. Given a corpus, broken down into individual works and within those works into individual sentences, you can query that corpus for all instances of a given word. The main unit of work is the `cps_corpus`, which is merely a collection of `cps_book` objects. These objects provide print methods to improve readability of the selections, always giving the author and the name of the work. 

## Basic functionality
There are two ways to use the `corpusr` package. First, if you are connected to the internet, you can communicate with the `gutenbergr` server to download and search books given an author's identification information. Second, if you would like to build your own corpus, you can use the `corpus_build` family of functions. 

### Interactive use
To work interactively, use the functions `guten_search_author` and `guten_search_works`. These are best used with information taken from `View(gutenbergr::gutenberg_metadata)`. 

### Corpus building
With the `guten_grab` function, you give a numeric code of a work and receive a `cps_corpus` with that work and its complete metadata. You can also build a corpus for all works by a certain author with `corpus_build_author`, and you can build a corpus with a list of book codes with `corpus_build_works`.

Once you have a corpus built, you can search it with `guten_search`. 
