---
title: "Corpusr: Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Corpusr: Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(corpusr)
```

## Introduction
The `corpusr` package shows you words in context. Given a corpus, broken down into individual works and within those works into individual sentences, you can query that corpus for all instances of a given word. The main unit of work is the `cps_corpus`, which is merely a collection of `cps_book` objects. These objects provide print methods to improve readability of the selections, always giving the author and the name of the work. 

## Basic functionality
There are two ways to use the `corpusr` package. First, if you are connected to the internet, you can communicate with the `gutenbergr` server to download and search books given an author's identification information. Second, if you would like to build your own corpus, you can use the `corpus_build` family of functions. 

### Interactive use
To work interactively, use the functions `guten_search_author` and `guten_search_works`. These are best used with information taken from `View(gutenbergr::gutenberg_metadata)`. 

### Corpus building
With the `guten_grab` function, you give a numeric code of a work and receive a `cps_corpus` with that work and its complete metadata. You can also build a corpus for all works by a certain author with `corpus_build_author`, and you can build a corpus with a list of book codes with `corpus_build_works`.

Once you have a corpus built, you can search it with `guten_search`. This function works by surrounding your argument with 'word limit' regular expression characters, so it is very flexible. For instance, you can search for both words and phrases, and you can use any of the typical regular expression toolkit you're used to. This flexibility allows you to search for verbs, for example, with a single search. You might be interested in the uses of the verb 'inure', so you would have a search that captures its typical English forms: `"inur(es|ed|ing)?"`. Or, you might be interested in certain collocations with "inure": has anyone ever been 'inured of' or 'inured on'? And so on. 
